---
title: "Activity_Analysis"
author: "Jeff Shelley"
date: "12/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr); library(caret); library(rpart.plot); library(caretEnsemble)
```

## Overview
Six participants are exercising with dumbbells in five different ways, and it's up to us to be able to recognize these five activities on a dataset where the activity isn't recorded.  We're going to split the training data into a train/test set, fitting two models and then stack them to see if we can get better results.  Then we'll load the validation set and see how we did.


```{r, echo=FALSE}
train.col.list = c("user_name",
              "roll_belt", "pitch_belt", "yaw_belt",
              "total_accel_belt",
              "gyros_belt_x", "gyros_belt_y","gyros_belt_z",
              "accel_belt_x","accel_belt_y","accel_belt_z",
              "magnet_belt_x","magnet_belt_y","magnet_belt_z",
              "roll_arm","pitch_arm","yaw_arm","total_accel_arm",
              "gyros_arm_x","gyros_arm_y","gyros_arm_z",
              "accel_arm_x","accel_arm_y","accel_arm_z",
              "magnet_arm_x","magnet_arm_y","magnet_arm_z",
              "roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
              "gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
              "accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
              "magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
              "roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm",
              "gyros_forearm_x","gyros_forearm_y","gyros_forearm_z",
              "accel_forearm_x","accel_forearm_y","accel_forearm_z",
              "magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
              "classe")
validate.col.list = c("user_name",
              "roll_belt", "pitch_belt", "yaw_belt",
              "total_accel_belt",
              "gyros_belt_x", "gyros_belt_y","gyros_belt_z",
              "accel_belt_x","accel_belt_y","accel_belt_z",
              "magnet_belt_x","magnet_belt_y","magnet_belt_z",
              "roll_arm","pitch_arm","yaw_arm","total_accel_arm",
              "gyros_arm_x","gyros_arm_y","gyros_arm_z",
              "accel_arm_x","accel_arm_y","accel_arm_z",
              "magnet_arm_x","magnet_arm_y","magnet_arm_z",
              "roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
              "gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
              "accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
              "magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
              "roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm",
              "gyros_forearm_x","gyros_forearm_y","gyros_forearm_z",
              "accel_forearm_x","accel_forearm_y","accel_forearm_z",
              "magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
              "problem_id")
```

First, let's set up our datasets, selecting a subset of columns (see Appendix for the columns used in the train/test and validation sets):
```{r}
data <- read.csv("data/pml-training.csv")
#classe is a factor variable
data$classe <- factor(data$classe)
#get rid of already-aggregated columns
data <- select(data,all_of(train.col.list))

#quick summary of the data we've selected from our Exploratory Data Analysis
str(data)


set.seed(654321)
inTrain <- createDataPartition(y=data$classe, p=0.75, list=FALSE)
train <- data[inTrain,]
dim(train)

test <- data[-inTrain,]
dim(test)

validation <- read.csv("data/pml-testing.csv")
#get rid of already-aggregated columns
validation <- select(validation, all_of(validate.col.list))
dim(validation)
```

Let's train with a couple different algorithms and then ensemble them!  
```{r}
set.seed(123654)
fit.1 <- train(
  classe ~ ., data = train,
  method = "rpart2", #note: rpart didn't find any case for D, but rpart2 does
  tuneLength=15)

set.seed(65432)
fit.2 <- train(classe ~ ., method="knn", data=train)
```
Make our Predictions with our two models and check their accuracies
```{r}
#now stack both the above
pred.1 <- predict(fit.1, newdata=test)
postResample(pred.1,test$classe)
pred.2 <- predict(fit.2, newdata=test)
postResample(pred.2,test$classe)
```

Combine the predictions of the two fitted models into a new data frame and train a third, stacked model against the test set:
```{r}
combined.df <- data.frame(pred1=pred.1, pred2=pred.2, classe=test$classe)

set.seed(65421)
fit.combined <- train(classe ~ pred1 + pred2, 
                      method="rf", data=combined.df)
pred.combined <- predict(fit.combined, newdata=combined.df)

postResample(pred.combined,test$classe)

```
So we've gone from fit.1 (75.04%) and fit.2 (91.70%) to 92.05% with them stacked!  We're feeling pretty good about this, let's take it to the validation set!

```{r}
pred.vd1 <- predict(fit.1, newdata=validation)
pred.vd2 <- predict(fit.2, newdata=validation)
combined.vddf <- data.frame(pred1=pred.vd1, pred2=pred.vd2, id=validation$problem_id)
pred.vdcombined <- predict(fit.combined, newdata=combined.vddf)
```

And a table of our final results:
```{r}
table(combined.vddf$id, pred.vdcombined)
```



## Appendix: Columns Used in data sets
```
train.col.list = c("user_name",
              "roll_belt", "pitch_belt", "yaw_belt",
              "total_accel_belt",
              "gyros_belt_x", "gyros_belt_y","gyros_belt_z",
              "accel_belt_x","accel_belt_y","accel_belt_z",
              "magnet_belt_x","magnet_belt_y","magnet_belt_z",
              "roll_arm","pitch_arm","yaw_arm","total_accel_arm",
              "gyros_arm_x","gyros_arm_y","gyros_arm_z",
              "accel_arm_x","accel_arm_y","accel_arm_z",
              "magnet_arm_x","magnet_arm_y","magnet_arm_z",
              "roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
              "gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
              "accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
              "magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
              "roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm",
              "gyros_forearm_x","gyros_forearm_y","gyros_forearm_z",
              "accel_forearm_x","accel_forearm_y","accel_forearm_z",
              "magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
              "classe")
validate.col.list = c("user_name",
              "roll_belt", "pitch_belt", "yaw_belt",
              "total_accel_belt",
              "gyros_belt_x", "gyros_belt_y","gyros_belt_z",
              "accel_belt_x","accel_belt_y","accel_belt_z",
              "magnet_belt_x","magnet_belt_y","magnet_belt_z",
              "roll_arm","pitch_arm","yaw_arm","total_accel_arm",
              "gyros_arm_x","gyros_arm_y","gyros_arm_z",
              "accel_arm_x","accel_arm_y","accel_arm_z",
              "magnet_arm_x","magnet_arm_y","magnet_arm_z",
              "roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
              "gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
              "accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
              "magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
              "roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm",
              "gyros_forearm_x","gyros_forearm_y","gyros_forearm_z",
              "accel_forearm_x","accel_forearm_y","accel_forearm_z",
              "magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
              "problem_id")
```
